\subsection{Frequency, Compression, and the Threshold for Collapse}

A key consequence of this picture is that the stability of light depends
on its transparency to the GCM. At low frequencies, LCM waves are
diffuse in absolute space and interact negligibly with the GCM flux.
They travel freely, their structure restored by the elastic properties
of the LCM.

When a light wave moves into a region of higher LCM compression, for example entering
the gravitational field of a planet or star, it blueshifts. That is, it's frequency increases
and the wave's spatial extent shrinks. The crests bunch
closer together, the oscillation occupies a smaller physical region, and
energy density rises. From the GCM perspective, the wave becomes a
denser and more coherent target.

This leads to a simple but profound rule:

\begin{quote}
\textbf{Blueshifting waves progressively increase their effective cross-section to the GCM flux.}
\end{quote}

At some point, the wave ceases to be transparent. The GCM begins
scattering off it strongly enough to cast a real shadow. The shadow
compresses the nearby LCM, which in turn constricts the wave. A
positive feedback loop takes over. Eventually the
compression crosses a critical threshold and the traveling wave loses the
ability to maintain its extended form. This is the onset of collapse.

Regions with high LCM density (such as near massive bodies or inside
planets) naturally compress incoming waves, increasing both their local
frequency and their opacity to the GCM. This explains why electron/positron pair
production is favored in such environments:

\begin{itemize}
\tightlist
\item
  \textbf{High local LCM density → higher local optical frequency}
\item
  \textbf{Higher optical frequency → increased GCM cross-section}
\item
  \textbf{Increased GCM opacity → stronger shadowing}
\item
  \textbf{Shadowing → collapse into standing waves}
\end{itemize}

Thus pair production is a natural consequence of the dual-medium
interaction. It occurs when a traveling wave becomes sufficiently
compact---and therefore sufficiently opaque to the GCM---that the only
stable configuration remaining is a pair of standing-wave particles with
complementary phase structures.
