The idea that Earth's radius may have changed over geological time has
appeared in many forms throughout the history of geophysics. Although it
sits outside the modern mainstream, it has been examined seriously by
researchers such as Samuel Warren Carey, who viewed expansion as a
possible unifying framework for continental drift, and Dr. James Maxlow,
who produced detailed reconstructions suggesting smaller ancient Earth
radii. A full discussion of the evidence and counter-arguments is beyond
the scope of this paper; readers interested in the broader debate are
encouraged to consult those primary works.

What matters for the present framework is not the historical argument itself, but the striking fact that it approaches the possibility of planetary growth from a completely different direction. Rather than beginning with geological interpretation, the model arrives at the question through the internal physics of the LCM–GCM interaction. Under conditions of sustained high compression in planetary interiors, the framework allows for ongoing proton formation as an irreversible channel through which energy can be converted into rest mass. This process, if present, would introduce slow mass increase over geological timescales, constrained by the available energy flux and local conditions rather than by surface geology alone.

One of the long-standing obstacles for expansion hypotheses has been the
absence of a physically grounded mechanism capable of driving sustained
planetary growth. 2MM does not claim to resolve the entire geological
debate, but it does offer a plausible mechanism where none previously
existed---providing a fresh context in which expansion models can be
reconsidered.
