\subsection{AGN Cores}\label{agn-cores}

In an infinite, steady-state universe, matter cannot simply accumulate.
If new standing waves can form in regions of high LCM compression, then
other regions must exist where standing waves can be dismantled and
their energy returned to the media. In 2MM, this role naturally falls to
the \textbf{innermost cores of active galactic nuclei (AGN)}---the only
environments dense enough to destabilize even neutron-like structures.

AGN can be viewed as layered systems:

\begin{itemize}
\tightlist
\item
  \textbf{Outer regions:} strong LCM gradients → frequent wave collapse
  → pair production and intermittent matter creation.
\item
  \textbf{Intermediate zones:} neutron-star--like compression →
  electrons and protons forced into neutron configurations.
\item
  \textbf{Innermost core:} LCM density exceeds the stability limit of
  the neutron itself.
\end{itemize}

At this deepest layer, 2MM proposes that the \textbf{neutron standing wave unravels}:

\begin{itemize}
\tightlist
\item
  the lead-mode and lag-mode components lose their phase-locked
  relationship,
\item
  the combined structure collapses into high-frequency LCM disturbances,
\item
  releasing energy back into the media in a neutral, charge-balanced
  way.
\end{itemize}

Because the neutron is globally neutral, its breakdown conserves charge
automatically. Any residual charged fragments are produced in matched
pairs or quickly reabsorbed by the surrounding plasma, ensuring that
\textbf{no net charge imbalance} is introduced.

Thus, matter destruction in 2MM is not the disappearance of isolated
protons, but a two-step sequence:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Extreme compression forces protons + electrons into neutrons}
  (charge conserved).
\item
  \textbf{Neutrons dissolve into high-frequency LCM waves} in cores
  where the standing-wave structure cannot survive.
\end{enumerate}

The resulting radiation escapes outward, where---under suitable
compression gradients---it can contribute to \textbf{new matter creation} via wave collapse.

In this picture, \textbf{AGN function as matter recyclers}: their outer
regions can create new standing waves, while their inner cores dismantle
old ones, maintaining balance in an infinite, ongoing universe.
