\section{Preface}

This document is an attempt to answer a question that has interested me
for many years: \emph{Is it possible to build a coherent, medium-based picture of physics that connects phenomena across scales, from particles to cosmology, using simple and intuitive principles?}

I approached this project not as a professional physicist but as an
amateur who has spent decades reading about science, puzzles, and
anomalies that invite deeper questioning. My interest has always leaned
toward the edges of what we understand: places where standard
explanations work mathematically but remain conceptually opaque, or
where observations fit uneasily into the prevailing frameworks.

The ideas presented here began with a few philosophical commitments:
that interactions should be local, that waves require a medium, and that
forces should have physical carriers rather than operating at a
distance. These assumptions are not universally accepted, and they are
not proposed as corrections to modern physics. They simply provided a
different starting point, one that led me to explore whether a
medium-based ontology might unify many disconnected pieces of physical
behavior.

This is not a finished theory. It is a conceptual model: speculative,
exploratory, and intentionally not defined mathematically. My hope is
that general readers find it accessible, and that professionals find it
interesting enough to question, critique, or even refine. Science
progresses not only through detailed calculation, but also through new
ways of seeing familiar things.
